### 2019-06-30 papers/journal/intro.tex
merged problem stat with the property section

In contrast, here we seek understand properties of vesicle traffic networks over all possible update rules, not just for a sampled subset.
%
Formal verification tools like model checkers, SAT,  SMT, and QBF solvers serve precisely this purpose. 
%
%Formal verification tools, such as model checkers
%\cite{clarke1996symbolic, biere2003bounded, clarke2008birth,
%	cimatti2000nusmv, holzmann1997model} and SAT solvers
%\cite{moskewicz2001chaff,een2004extensible} serve precisely this
%purpose.
%%
The tools model the computation symbolically for the graphs of finite size and can make a prediction about the hypothesis for all possible states of the system without explicit enumeration of rules. 
%
In a recent work \cite{shukla2017discovering}, we employed the model checker CBMC \cite{ckl2004} meant to analyze ANSI C programs for studying this problem. 
%
VTSs and SNARE regulations were modeled as C code using arrays to represent graphs and Boolean tables. 
%
The graph connectivity question for different variations of SNARE regulations was modeled as logical assertions to be checked by the model checker. 
%
Although the performance was satisfactory the encoding was non-optimal and hence scalability of the method was insufficient (the tool was able to handle the system with just 8 compartments).
%
Another interesting question of property involving quantifiers (was termed as sufficient condition in \cite{shukla2017discovering}; in this paper, we call it universal condition) was partly handled using sampling and left open as future work.
%
In this paper, we present both the SMT and QBF encodings to address these two problems.
%
%These tools are central to formal verification methods in computer science, and are increasingly being applied to understand biological systems, including gene regulatory networks [16-18]. 
%\subsection{Previous work and our Contribution}
%
%The following are the contributions of this work:
%\begin{itemize}
%	\item We have added novel encoding reachability and 3-4 connectivity
%	\item Direct encoding into the SAT solver
%	\item A user friendly and scalable tool based on well known SMT solver Z3
%\end{itemize}

@@

%
The rest of the paper is organized as follows. 
%
In Section~\ref{sec:bio}, we present the biological background and details to understand the VTS. 
%
In Section~\ref{sec:prelim}, we present the notations we use in this paper. 
%
In Section~\ref{sec:model}, we present a formal model of VTSs and in Section~\ref{sec:property} the properties of interest.
%
In Section~\ref{sec:problem}, we present the formal problem statements of different search problems of VTS.
%
In Section~\ref{sec:encoding}, we present the encodings of corresponding search problems. 
%
In Section~\ref{sec:experiments}, we present our implementation and experimental results. 
%
In Section~\ref{sec:related}, we discuss related work and conclude in Section~\ref{sec:conclusion}.
%
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
~

>>>>

In contrast, here we seek understand properties of vesicle traffic networks over all possible update rules, not just for a sampled subset.
%
Formal verification tools like model checkers, SAT,  SMT, and QBF solvers serve precisely this purpose. 
%
%Formal verification tools, such as model checkers
%\cite{clarke1996symbolic, biere2003bounded, clarke2008birth,
%	cimatti2000nusmv, holzmann1997model} and SAT solvers
%\cite{moskewicz2001chaff,een2004extensible} serve precisely this
%purpose.
%%
The tools model the computation symbolically for the graphs of finite size and can make a prediction about the hypothesis for all possible states of the system without explicit enumeration of rules. 
%
In a recent work \cite{shukla2017discovering}, we employed the model checker CBMC \cite{ckl2004} meant to analyze ANSI C programs for studying this problem. 
%
VTSs and SNARE regulations were modeled as C code using arrays to represent graphs and Boolean tables. 
%
The graph connectivity question for different variations of SNARE regulations was modeled as logical assertions to be checked by the model checker. 
%
Although the performance was satisfactory the encoding was non-optimal and hence scalability of the method was insufficient (the tool was able to handle the system with just 8 compartments).
%
Another interesting question of property involving quantifiers (was termed as sufficient condition in \cite{shukla2017discovering}; in this paper, we call it universal condition) was partly handled using sampling and left open.
%
In this paper, we present both the SMT and QBF encodings to address these two problems.
%
%These tools are central to formal verification methods in computer science, and are increasingly being applied to understand biological systems, including gene regulatory networks [16-18]. 
%\subsection{Previous work and our Contribution}
%
%The following are the contributions of this work:
%\begin{itemize}
%	\item We have added novel encoding reachability and 3-4 connectivity
%	\item Direct encoding into the SAT solver
%	\item A user friendly and scalable tool based on well known SMT solver Z3
%\end{itemize}

@@

%
The rest of the paper is organized as follows. 
%
In Section~\ref{sec:bio}, we present the biological background and details to understand the VTS. 
%
In Section~\ref{sec:prelim}, we present the notations we use in this paper. 
%
In Section~\ref{sec:model}, we present a formal model of VTSs and in section~\ref{sec:problem}, we present the property of interest with their formal problem statements of different search problems of VTS.
%
In Section~\ref{sec:encoding}, we present the encodings of corresponding search problems. 
%
In Section~\ref{sec:experiments}, we present our implementation and experimental results. 
%
In Section~\ref{sec:related}, we discuss related work and conclude in Section~\ref{sec:conclusion}.
%
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
~

### 2019-06-30 papers/journal/prelim.tex
Update first iteration on Review
merged problem stat with the property section

\subsubsection{Solvers}
In this paper, we will encode several problems into finding satisfying
solutions of a Boolean formula or QBFs.
%
We will use SAT, SMT and QBF solvers to get the satisfying assignments.
%
The modern solvers are the result of several decades of research
and implement sophisticated algorithms.
%
A SAT solver typically implements conflict-driven clause learning (CDCL)~\cite{biere2009conflict}.
%1
The QBF solver adds support of quantifiers by adding more rules for quantifiers~\cite{buning2009theory}.

%
% Due to the combinatorial nature of the graphs, the search space is huge
% and often hard to enumerate naively.
%
% We need sophisticated solvers

>>>>

\subsubsection{Solvers}
In this paper, we will encode several problems into finding satisfying solutions of a Boolean formula or QBFs.
%
We will use SAT, SMT and QBF solvers to get the satisfying assignments.
%
The modern solvers are the result of several decades of research
and implement sophisticated algorithms.
%
%
A SAT solver typically implements conflict-driven clause learning (CDCL)~\cite{biere2009conflict}.
%1
SMT solvers~\cite{nieuwenhuis2006solving, barrett2018satisfiability} in addition adds supports for some first order theories like arithmetic, arrays etc.
%

The QBF solver adds support of quantifiers by adding more rules for quantifiers~\cite{buning2009theory}.

We also use bounded model checkers that implements Bounded model checking (BMC) technique~\cite{biere1999symbolic1, biere2003bounded} to performs a search for the violation of the property in the program executions whose length is bounded by some fixed positive integer value $w$ (\textit{unwinding depth}). 
%
% Due to the combinatorial nature of the graphs, the search space is huge
% and often hard to enumerate naively.
%
% We need sophisticated solvers

### 2019-06-30 papers/journal/problem.tex
Update first iteration on Review
merged problem stat with the property section

\noindent This section will present the different search problems of VTS.
%
We categorize the search problem (Fig.~\ref{fig:vts-search}) into three different categories, search problem for the existential condition, search problem for the universal condition, and performing synthesis on the incomplete VTS. 
%
We will present two encodings for existential condition and one each for universal condition and synthesis.

@@

\subsection{Existential condition}
\noindent We employ two techniques to examine the existential condition. 
%
First, using the bounded model checking (BMC) technique~\cite{biere1999symbolic1, biere2003bounded}.
%
%It  represents the executions of a program symbolically and 
%
%The Bounded model checking technique
BMC is an automatic technique that performs a search for the violation of the property in the program executions whose length is bounded by some fixed positive integer value $w$ (\textit{unwinding depth}). 
%
The bounded execution is represented symbolically as a propositional formula and solved using a SAT solver.
%
In case there is no violation for a particular depth, one can progressively increase the depth until either property is violated, the problem becomes intractable, or an upper bound of the system is reached, indicating the satisfiability of the property.
%
For our case, the unwinding depth is the length of the path of the underline graph.
%
%
For the experimentation, we restrict the check only to the elementary cycles; cycles of length one. 
%
This restriction puts an upper limit of the number of nodes on the length of the searched path. 
%
%
%and the upper bound is the number of nodes $N$.

Second, using constraint solving, where the rules are represented as a constraint over the variables defined in the system.
%
The search problem is to find an assignment of variables that satisfies all the specified constraints.
%
We have used a form of this technique known as SMT~\cite{nieuwenhuis2006solving, barrett2018satisfiability}.
%
The SMT is a generalization of Boolean satisfiability problem with domain-specific reasoning, for example, support for uninterpreted functions, equality, arithmetic and data structures like arrays, bit-vectors, etc. 
%
%It includes formulas with variables from other domains, including real numbers, integers, and support for various data structures such as lists, arrays, bit vectors.
% is satisfiability of formulas with respect to some background theory.
%
Throughout the paper, we will address these two subcategories as $\fbmc$ and $\fsmt$ search problems.
%The existential condition search problem is represented as a bounded model checking and constraint solving. 
%
%Based upon earlier discussion, we need to answer the following search question among VTSs.
%
Here, we present both the $\fbmc$ and $\fsmt$ search problem for the existential condition.

@@

%\noindent
%The notion of stability and connectivity varies slightly for BMC encoding. 
%%In this subsection, we will present the concept of stability and connectivity for BMC encoding. 
%%
%We call a graph $G$ {\em stable} if for each $(n,M',n') \in \edges$ and $m \in M'$,
%there exists a $m$-path in G with $n_1 = n'$ and $n_\ell = n$.
%%
%%
%We call a graph $G$ {\em connected} if for each $n,n' \in \nodes$,
%$n'$ there is a path from $n$ to $n'$ in G.
%
%We call a graph $G$ $k$-connected if for each $\edges' \subseteq \edges$ and $|\edges'| < k$,
%VTS $(\nodes,\mols,\edges-E',\nlabel,\pairs,\edgef,\nodef)$ is connected.
%%	
%We use $m$-${path}_i^j$ to denote an $m$-path between source node i and target node is j. 
%
%
%Creating a formula to search a path of length N-1 for each molecule m on every edge E will be extremely large and harder for the solvers to handle.  
%%
%Even if there exist shorter paths we have to create the formula of maximum size N-1. 
%% 
%To resolve these problems we will use a technique called bounded model checking, where the depth of the path can be specified and consequently increased or decreased.  
%

\subsection{Universal condition}
%Here we will present the QBF.
\noindent We will model the search problem for the universal condition as a QBF~\cite{buning2009theory, benedetti2008qbf}, and the resultant formula is solved using a QBF solver.
%
The QBF is a generalization of the Boolean satisfiability problem in which every variable is quantified using either existential or universal quantifier.
%
The QBFs enables potentially more succinct encodings than SAT~\cite{jussila2007compressing} although much more difficult~\cite{savitch1970relationships, stockmeyer1973word} to solve.

@@

We will do synthesis against the property that the VTS
is stable and 3-connected.
%
% The property is designed to balance the search space such that the synthesis procedure does not
% succeed with simply adding too many edges. 
%
%\begin{align*}
%\texttt{Property} =  \texttt{Stability} \land \texttt{Connected}(3) 
%%\land \texttt{DisConnected}(4)
%\end{align*}
This property was proposed in~\cite{shukla2017discovering}.
%
However, the biological relevance of the property is debatable and open for change.
%
% Our tool is easily modifiable to support any other property that may be deemed 
% interesting by the biologists.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

>>>>

\noindent %\improve{The weakest part of the paper; Need a serious rewrite!}
We analyze the model against our desired properties.
%
%Since in biology there are no declared specifications or properties, we conjecture some properties that appear to be 
%true in  relevant inall the known biological systems.
%
In this paper, we choose graph connectedness to be a key property.

We categorize the search problem (Fig.~\ref{fig:vts-search}) into three different categories, search problem for the existential condition, search problem for the universal condition, and performing synthesis on the incomplete VTS. 
%
We will present two encodings for existential condition and one each for universal condition and synthesis.

@@

%
We will also seek two kinds of questions give the property.
% %
% In this paper,
% our conjecture relates graph-connectedness to a particular variation of SNARE pairing and
% regulation rules of the VTS.
% %
% We are interested in the following two type of properties of the transport graph:
\begin{enumerate}
\item {\em Existential condition}:
if there is a VTS that satisfies the connectedness property.

\item {\em Universal condition}:
  if there is a connectedness property that is satisfied by each valid VTS.
\end{enumerate}
The goal is to find the minimum graph connectivity properties for each of the condition.
% In our program, the rules VTS needs to follow 
% %
% are represented as constrained over the transport graphs (G), label on the nodes and edges and activity of those labels (determined by a Boolean function f). 
% %
% Each of these constrained functions is defined to be TRUE for a G and f: if and only if the corresponding condition holds for the given G and f. 
% %
% We use these functions to define the assertion that characterizes the corresponding condition to be checked. 
% %
% We will show in later sections how we encode these Boolean functions. 
\subsubsection{Search for existential condition} 
%
% To ensure that k-connectedness is an existential condition,
%
We need to find a $k$-connected graph that satisfies all the rules (and constraints) of the system. 
%
We also aim to find the minimum $k$. 
%
%Let's fix our conjecture: ``k-connectedness is a necessary condition for the system regulated by a Boolean function on the node and SNARE-SNARE inhibition on the edge". 
%
To find a k-connected graph we specify our property as a query for the existence
of a model for a conjunction of the formula representing stability, fusion and specific
graph connectivity constraint. 
%
% \begin{align*}
%   &\texttt{LabeledVTS} = \texttt{Stability} \, \land \, \texttt{Fusion} \land \, \texttt{MinConnectivity(k)}
%  \tag{E}\label{eq:existcond}
% \end{align*}
%
We will start with the value of $k = 1$ and check for the satisfiability of the formula.
%
In case the formula is satisfiable the procedure terminates and we report the value of $k$. 
%
If the formula is not satisfiable we increment the $k$ by $1$ and repeat the same procedure. 
%
In this way, we ensure that the reported $k$ is the minimum connectivity of the VTS that
satisfies the existential condition.
%
% Used E as edge and L_e L_n as label for node and the edge,

 We employ two techniques to examine the existential condition. 
%
First, using the bounded model checking (BMC) technique~\cite{biere1999symbolic1, biere2003bounded}.
%
%It  represents the executions of a program symbolically and 
%
%The Bounded model checking technique
%BMC is an automatic technique that performs a search for the violation of the property in the program executions whose length is bounded by some fixed positive integer value $w$ (\textit{unwinding depth}). 
%%
%The bounded execution is represented symbolically as a propositional formula and solved using a SAT solver.
%%
%In case there is no violation for a particular depth, one can progressively increase the depth until either property is violated, the problem becomes intractable, or an upper bound of the system is reached, indicating the satisfiability of the property.
%%
%For our case, the unwinding depth is total number of nodes in the graph.
%the maximum length of the path of the underline graph.
%
%
%For the experimentation, we restrict the check only to the elementary cycles; cycles of length one. 
%%
%This restriction puts an upper limit of the number of nodes on the length of the searched path. 
%
%
%and the upper bound is the number of nodes $N$.
Second, using constraint solving, where the rules are represented as a constraint over the variables defined in the system.
%
The search problem is to find an assignment of variables that satisfies all the specified constraints.
%
We have used SMT solving.

Throughout the paper, we will address these two subcategories as $\fbmc$ and $\fsmt$ search problems.
%The existential condition search problem is represented as a bounded model checking and constraint solving. 
%
%Based upon earlier discussion, we need to answer the following search question among VTSs.
%
Here, we present both the $\fbmc$ and $\fsmt$ search problem for the existential condition.

@@

\subsubsection{Search for universal condition}
%
% Similarly, for the conjecture that k-connectedness is a universal condition for the VTS,
For some $k>0$,
we have to ensure that every graph of $k$-connected structure satisfies all
the rules of the system.
%
We also aim to find the minimum $k$. 
% We also have to ensure that the particular k is the least one. 
%
This condition can be specified as: for every k-connected graph, there exists a satisfiable assignment following the rules of the system. 
%
% \begin{align*}
%   & \forall G \, \texttt{Connectivity(k)} \implies  \exists
%                         f,p: \texttt{Stability} \, \land \, \texttt{Fusion}  
%   \tag{U}\label{eq:univcond}
% \end{align*}
%
We will use the same procedure as the existential condition to ensure the k to be the least one.
%
The formula for the problem has quantifier alternation. 
%
Most SAT solvers can check (at least efficiently) only quantifier-free formulas.
%
We need a QBF solver for such queries.
%
We will model the search problem for the universal condition as a QBF~\cite{buning2009theory, benedetti2008qbf}, and the resultant formula is solved using a QBF solver.
%
The QBF is a generalization of the Boolean satisfiability problem in which every variable is quantified using either existential or universal quantifier.
%
The QBFs enables potentially more succinct encodings than SAT~\cite{jussila2007compressing} although much more difficult~\cite{savitch1970relationships, stockmeyer1973word} to solve.

@@

We will do synthesis against the property that the VTS
is stable and 3-connected.
%
% The property is designed to balance the search space such that the synthesis procedure does not
% succeed with simply adding too many edges. 
%
%\begin{align*}
%\texttt{Property} =  \texttt{Stability} \land \texttt{Connected}(3) 
%%\land \texttt{DisConnected}(4)
%\end{align*}
This property was proposed in~\cite{shukla2017discovering}.
%
However, the biological relevance of the property is debatable and open for change.
%

%To handle this challenge, we used a combination of nondeterminism and Boolean enumeration at the C-source level to eliminate quantifiers, as explained further below.
%
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
